A Triple Origin for Twin Blue Stragglers in Compact Binaries

### 2018-12-02 mn.tex
Add files via upload

%%
%% Beginning of file 'sample62.tex'
%%
%% Modified 2018 January
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.2 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass{aastex62}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex62}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Preferred style for Research Notes which are by design 
%%                lacking an abstract and brief. DO NOT use \begin{abstract}
%%                and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{January 1, 2018}
\revised{January 7, 2018}
\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Compact Binaries of BS Twins from Stellar Triples}
\shortauthors{Leigh \& Portegies Zwart}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

@@

\title{A Triple Origin for Twin Blue Stragglers in Compact Binaries}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.2 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.2 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Nathan W. C. Leigh}
\email{nleigh@amnh.org}

\author{Nathan W. C. Leigh}
\affil{American Museum of Natural History \\
Department of Astrophysics \\
79th Street at Central Park West \\
New York, NY 10024-5192, USA}
\affil{Stony Brook University \\
Department of Physics and Astronomy\\
Stony Brook, NY 11794-3800, USA}
\affil{Departamento de Astronom\'ia \\ 
Facultad de Ciencias F\'isicas y Matem\'aticas \\ 
Universidad de Concepci\'on \\ 
Concepci\'on, Chile}

\author{Simon Portegies Zwart}
\affiliation{Leiden Observatory \\
Leiden University \\
PO Box 9513, 2300 RA \\
Leiden, the Netherlands}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

\begin{abstract}
In this Letter, we propose a potential formation mechanism for twin blue stragglers in compact binaries that involves mass transfer from an evolved outer tertiary companion on to the inner binary via a circumbinary disk.  We apply our hypothetical scenario to the observed double blue straggler system Binary 7782 in the old open cluster NGC 188, and show that its observed properties are easily and even naturally reproduced within the context our proposed model.  Within the context of the hypothetical formation mechanism proposed here, the presented work predicts the following properties for the post-mass transfer double BS tertiary:  (1)  For the outer tertiary orbit, the orbital period should lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days, assuming initial masses for the inner binary components of m$_{\rm 1} =$ 1.1 M$_{\odot}$ and m$_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer tertiary mass of m$_{\rm 3} =$ 1.4 M$_{\odot}$. (2)  Larger final WD masses, and hence core masses for the donor at the time of mass transfer, should correspond to larger final outer orbital periods for the tertiary. (3)  For the inner binary, the rotational axes of the BSs should be aligned with each other and the orbital plane of the outer tertiary WD. (4)  The BSs in the inner binary should have roughly equal masses, independent of their initial masses.  This predicts that the initially lower mass MS star should accrete the most, and should hence be polluted more significantly by any accreted material.  This could be observable in the surface layers of a radiative star (i.e., He, C and O if the donor is an RGB star, and/or s-process elements if the donor is an AGB star). (5) Twin BSs in compact binaries formed from the proposed mechanism should be more frequent in younger clusters with ages $\lesssim$ 4-6 Gyr, since the donor will have  radiative envelope.

@@

\keywords{stars: blue stragglers -- binaries: general -- globular clusters: general -- scattering}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{intro}

Blue straggler (BS) stars appear brighter and bluer than the main-sequence turn-off (MSTO) in a cluster colour-magnitude diagram (CMD) \citep[e.g.][]{simunovic14,simunovic16}.  Two primary channels for BS formation have been proposed; mass transfer from an evolved donor on to a main-sequence star in a binary star system \citep[e.g.][]{mccrea64,knigge09,mathieu09,leigh11,geller11,geller12,gosnell14,gosnell15}, and direct stellar collisions involving main-sequence stars mediated via direct interactions involving binaries \citep[e.g.][]{hills75,shara97,leigh07,leigh11,leigh13,hypki13}.  Other possible, albeit related, formation mechanisms include mergers of compact MS-MS binaries, and mergers of the inner MS-MS binaries of hierarchical triple star systems induced by Lidov-Kozai oscillations coupled with tidal damping \citep[e.g.][]{perets09}.

@@

Figure~\ref{fig:fig1} shows the parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial component masses of m$_{\rm 1} =$ 1.1 M$_{\rm \odot}$ and m$_{\rm 2} =$ 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3} =$ 1.4 M$_{\rm \odot}$ for the outer tertiary.  We compare the circularization radius to the semi-major axis of the inner binary, for which we require a$_{\rm c} >$ a$_{\rm in}$, after folding in all constraints from the requirements for dynamical stability, an outer tertiary that is Roche lobe-filling, and a dynamically hard outer tertiary orbit.  Note that the range of plotted orbital periods P$_{\rm in}$ corresponding to a contact state for the inner binary, assuming R$_{\rm 1} =$ R$_{\rm 2} =$ 1 R$_{\odot}$, lies outside the range of plotted values for P$_{\rm in}$, since it does not contribute significantly to constraining the outer orbital properties of a hypothesized outer WD tertiary.  The thick horizontal solid red line shows the allowed range of outer semi-major axes, after folding in all of the aforementioned criteria.  As is clear, this makes a relatively narrow prediction for the allowed ranges of outer tertiary orbits, namely 2.2 $\times$ 10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days, for our assumed final donor mass.

\begin{figure}[ht!]
\plotone{fig1.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane allowed for the hypothetical outer tertiary orbit of Binary 7782.  The solid diagonal black lines show the period corresponding to the circularization radius a$_{\rm c}$ for the mass transfer stream coming from the outer tertiary, for different values of the mass ratio, namely q $=$ 0.1, 0.7, 1, 2 and 10.  We assume initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ is computed for the outer tertiary according to our assumed mass ratio (with our fiducial case corresponding to q $=$ 0.7.  We assume completely conservative mass transfer for this exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal black line shows a rough criterion for dynamical stability in the triple, approximately following \citet{mardling99} (i.e., a$_{\rm in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical stability in equal-mass co-planar triples).  The vertical solid red lines show the outer orbital periods corresponding to the hard-soft boundary assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km s$^{-1}$.  The vertical dashed black lines show the maximum outer orbital period P$_{\rm out}$ for which the outer tertiary companion is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$ 200 R$_{\odot}$ (CHANGE, POSSIBLY USING MESA CALCULATIONS?!).  The horizontal dashed red line shows the observed orbital period for Binary 7782, using its observed orbital period and our assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line shows the parameter space for P$_{\rm out}$ allowed after considering all of the aforementioned criteria.
\label{fig:fig1}}
\end{figure}

@@

\item For the predicted outer tertiary orbit, the present-day semi-major axis should lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$ 1100 days, assuming initial masses for the inner binary components of m$_{\rm 1} =$ 1.1 M$_{\odot}$ and m$_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer tertiary mass of m$_{\rm 3} =$ 1.4 M$_{\odot}$.

@@

\acknowledgments

N.W.C.L. acknowledges support from a Kalbfleisch Fellowship at the American Museum of Natural History.

\begin{thebibliography}{}

%\bibitem[\protect\citeauthoryear{Anderson et al.}{2008}]{anderson08}
% Anderson J.,  Sarajedini A., Bedin L. R., King I. R., Piotto G.,
%  Reid I. N., Siegel M., Majewski S. R., Paust N. E. Q., Aparicio A.,
%  Milone A. P., Chaboyer B., Rosenberg A. 2008, AJ, 135, 2055
%\bibitem[\protect\citeauthoryear{Artymowicz et al.}{1993}]{artymowicz93} Artymowicz P., Lin D. N. C., Wampler E. J. 1993, ApJ, 409, 592 

@@

%\bibitem[\protect\citeauthoryear{Bedin et al.}{2010}]{bedin10} Bedin L. R., Cassisi S., Castelli F., Piotto G., Anderson J., Salaris M., Momany Y., Pietrinferni A. 2010, MNRAS 357, 1038

@@

%\bibitem[\protect\citeauthoryear{Bruzual \& Charlot}{2003}]{bruzual03} Bruzual G., Charlot S. 2003, MNRAS, 344, 1000 
%\bibitem[\protect\citeauthoryear{Casassus et al.}{2013}]{casassus13} Casassus S., Hales A., de Gregorio I., Dent B., Belloche A., G\"usten R., M\'enard F., Hughes A. M., Wilner D., Salinas V. 2013, A\&A, 553, 64 
%\bibitem[\protect\citeauthoryear{Ceillier et al.}{2017}]{ceillier17} Ceillier T., Tayar J., Mathur S., Salabert D., Garc\'ia R. A., Stello D., Pinsonneault M. H., van Sanders J., Beck P. G., Bloemen S. 2017, A\& A, 605, 111 

@@

%\bibitem[\protect\citeauthoryear{Dotter et al.}{2010}]{dotter10} Dotter, A., Sarajedini A., Anderson J., Aparicio A., Bedin L. R., Chaboyer B., Majewski S., Marin-Franch A., Milone A., Paust N., Piotto G., Reid N., Rosenberg A., Siegel M. 2010, ApJ 708, 698 
%\bibitem[\protect\citeauthoryear{Dalessandro et al.}{2013}]{dalessandro13} Dalessandro E., Ferraro F. R., Massari D., Lanzoni B., Miocchi R. P., Beccari G., Bellini A., Sills A., Sigurdsson S., Mucciarelli A., Lovisi L. 2013, ApJ, 778, 135
%\bibitem[\protect\citeauthoryear{Eggleton}{1983}]{eggleton83} Eggleton P. P. 1983, ApJ, 268, 368 
%\bibitem[\protect\citeauthoryear{Flock et al.}{2015}]{flock15} Flock M., Ruge J. P., Dzyurkevich N., Henning Th., Klahr H., Wolf S. 2015, A\&A, 574, 68
%\bibitem[\protect\citeauthoryear{Ferraro et al.}{1997}]{ferraro97}
%  Ferraro F. R., Paltrinieri B., Fusi Pecci F., Cacciari C., Dorman
%  B., Rood R. T., Buonanno R., Corsi C. E., Burgarella D., Laget
%  M. 1997, A\&A, 324, 915
%\bibitem[\protect\citeauthoryear{Ferraro et al.}{1999}]{ferraro99}
%  Ferraro F. R., Paltrinieri B., Rood R. T., Dorman B. 1999, ApJ, 522,
%  983
%\bibitem[\protect\citeauthoryear{Ferraro et al.}{2004}]{ferraro04}
%  Ferraro F. R., Beccari G., Rood, R. T., Bellazzini M., Sills A.,
%  Sabbi E. 2004, ApJ, 603, 127
%\bibitem[\protect\citeauthoryear{Ferraro, Valenti \& Origlia}{2006}]{ferraro06} Ferraro F. R., Valenti E., Origlia L. 2006, ApJ, 649, 243

>>>>

%mm4 comments/revisions on v8 labeled with %mm4
%mm5 revisions on v10 labeled with %mm5
%
\documentclass[useAMS,usenatbib]{mnras}
\topmargin -1cm
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{amssymb}\usepackage{hyperref}
\usepackage{amsmath,alltt}                                                                                                                     
\usepackage{multirow}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{times}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{supertabular}

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$<$}}}
\newcommand\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$>$}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ba}{\begin{eqnarray}}
%\newcommand{\baa}{\begin{align}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ea}{\end{eqnarray}}
%\newcommand{\eaa}{\end{align}}

%\newcommand{\mnras}{\textrm{MNRAS}}
%\newcommand{\apj}{\textrm{ApJ}}
%\newcommand{\aap}{\textrm{A\&A}}
%\newcommand{\apjl}{\textrm{ApJ}}
%\newcommand{\prd}{\textrm{Phys Rev D}}

\newcommand{\comment}{\bf}
\newcommand{\pont}{{\,^\ast\!}R\,R}
\newcommand{\red}{\color{red}}
\newcommand{\blue}{\color{blue}}
\newcommand{\green}{\color{green}}
\newcommand{\black}{\color{black}}
\newcommand{\Scri}{{\mathcal I}}
\newcommand{\re}{{\textrm{real}}}
\newcommand{\eff}{{\textrm{eff}}}
\newcommand{\p}{{\textrm{pole}}}
\newcommand{\lf}{{\textrm{log-fac}}}
\newcommand{\SMBH}{\bullet}
\newcommand{\SCO}{\star}
\newcommand{\GW}{{\mbox{\tiny GW}}}
\newcommand{\vac}{{\mbox{\tiny vac}}}
\newcommand{\SPA}{{\mbox{\tiny SPA}}}
\newcommand{\Tot}{{\mbox{\tiny tot}}}
\newcommand{\Bondi}{{\mbox{\tiny Bondi}}}
\newcommand{\Ho}{{\mbox{\tiny Hor}}}
\newcommand{\Pa}{{\mbox{\tiny Pade}}}
\newcommand{\pd}{{\mbox{\tiny PD}}}
\newcommand{\Newt}{{\mbox{\tiny Newt}}}
\newcommand{\Pert}{{\mbox{\tiny Pert}}}
\newcommand{\Ta}{{\mbox{\tiny Tay.}}}
\newcommand{\lfac}{{\mbox{\tiny log-fac}}}
\newcommand{\NS}{{\mbox{\tiny NS}}}
\newcommand{\Edd}{{\mbox{\tiny Edd}}}
\newcommand{\Sp}{{\mbox{\tiny S}}}
\newcommand{\ISCO}{{\mbox{\tiny ISCO}}}
\newcommand{\initial}{{\mbox{\tiny Initial}}}
\newcommand{\final}{{\mbox{\tiny Final}}}
\newcommand{\mock}{{\mbox{\tiny mock}}}
\newcommand{\ny}[1]{\textcolor{red}{\it{\textbf{ny: #1}}} }
\newcommand{\sah}[1]{\textcolor{magenta}{\it{\textbf{sah: #1}}} }
\newcommand{\bk}[1]{\textcolor{blue}{\it{\textbf{bk: #1}}} }
\newcommand{\OrbitalFreq}{\Omega}
\newcommand{\Sm}{{\mbox{\tiny S}}}
\newcommand{\La}{{\mbox{\tiny L}}}
\newcommand{\EOB}{{\mbox{\tiny EOB}}}
\newcommand{\Msun}{\,{\rm M_\odot}}
\newcommand{\yr}{\,{\rm yr}}
\newcommand{\cm}{\,{\rm cm}}
\newcommand{\gm}{\,{\rm g}}
\newcommand{\I}{\mathcal{J}}
\newcommand{\D}{\mathrm{d}}
\newcommand{\orb}{\mathrm{orb}}
\newcommand{\R}{{\bar{r}}}
\newcommand{\Ls}{\ell}
\newcommand{\gas}{\mathrm{gas}}
\newcommand{\gap}{\mathrm{gap}}
\newcommand{\dL}{d_{\rm L}}

\title[Compact Binaries of BS Twins from Stellar Triples]{A Triple Origin for Twin Blue Stragglers in Compact Binaries}

\author[N. W. C. Leigh \& S. Portegies Zwart]{N. W. C. Leigh$^{1,2,3}$, S. Portegies Zwart$^{4}$ \\
$^{1}$Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA\\
$^{2}$Department of Physics and Astronomy, Stony Brook University, Stony Brook, NY 11794-3800, USA\\
$^{3}$Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Avenue, New York, NY 10010, USA\\
$^{4}$Leiden Observatory, }

@@

\date{Accepted. Received; in original form}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2008}

\maketitle

\label{firstpage}

\begin{abstract}
In this Letter, we propose a potential formation mechanism for twin blue stragglers in compact binaries that involves mass transfer from an evolved outer tertiary companion on to the inner binary via a circumbinary disk.  We apply our hypothetical scenario to the observed double blue straggler system Binary 7782 in the old open cluster NGC 188, and show that its observed properties are easily and even naturally reproduced within the context our proposed model.  Within the context of the hypothetical formation mechanism proposed here, the presented work predicts the following properties for the post-mass transfer double BS tertiary:  (1)  For the outer tertiary orbit, the semi-major axis should lie between 2 AU $\lesssim$ a$_{\rm out}$ $\lesssim$ 5 AU, assuming initial masses for the inner binary components of m$_{\rm 1} =$ 1.1 M$_{\odot}$ and m$_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer tertiary mass of m$_{\rm 3} =$ 1.4 M$_{\odot}$. (2)  Larger final WD masses, and hence core masses for the donor at the time of mass transfer, should correspond to larger final outer orbital periods for the tertiary. (3)  For the inner binary, the rotational axes of the BSs should be aligned with each other and the orbital plane of the outer tertiary WD. (4)  The BSs in the inner binary should have roughly equal masses, independent of their initial masses.  This predicts that the initially lower mass MS star should accrete the most, and should hence be polluted more significantly by any accreted material.  This could be observable in the surface layers of a radiative star (i.e., He, C and O if the donor is an RGB star, and/or s-process elements if the donor is an AGB star). (5) Twin BSs in compact binaries formed from the proposed mechanism should be more frequent in younger clusters with ages $\lesssim$ 4-6 Gyr, since the donor will have  radiative envelope.

@@

\begin{keywords}
stars: blue stragglers -- binaries: general -- globular clusters: general -- scattering. 
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Blue straggler (BS) stars appear brighter and bluer than the main-sequence turn-off (MSTO) in a cluster colour-magnitude diagram (CMD).  Two primary channels for BS formation have been proposed; mass transfer from an evolved donor on to a main-sequence star in a binary star system \citep[e.g.][]{mccrea64,knigge09,leigh11,gosnell14,gosnell15}, and direct stellar collisions involving main-sequence stars mediated via direct interactions involving binaries \citep[e.g.][]{hills75,shara97,leigh13,hypki13}.  Other possible, albeit related, formation mechanisms include mergers of compact MS-MS binaries, and mergers of the inner MS-MS binaries of hierarchical triple star systems induced by Lidov-Kozai oscillations coupled with tidal damping \citep[e.g.][]{perets09}.

@@

Figure~\ref{fig:fig1} shows the parameter space in the a$_{\rm out}$-a$_{\rm in}$-plane for Binary 7782.  We assume initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ = 1.4 M$_{\rm \odot}$ for the outer tertiary.  This choice is guided by the current turn-off mass of 1.1 M$_{\odot}$ in NGC 188 \citep{mathieu09} and a maximum BS age of a few Gyr \citep[e.g.][]{sills99,sills01}.  After comparing the circularization radius to the semi-major axis of the inner binary and folding in all constraints from the requirements for dynamical stability, an outer tertiary that is Roche lobe-filling, and a dynamically hard outer tertiary orbit.  The thick horizontal solid red line shows the allowed range of outer semi-major axes allowed, after folding in all of the aforementioned criteria.  As is clear, this makes a relatively narrow prediction for the allowed ranges of outer tertiary orbits, namely 2 AU $\le$ a$_{\rm out} \le$ 5 AU.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{fig1.eps}
\end{center}
\caption[Parameter space in the a$_{\rm out}$-a$_{\rm in}$-plane allowed for the hypothetical outer tertiary orbit of Binary 7782.]{Parameter space in the a$_{\rm out}$-a$_{\rm in}$-plane allowed for the hypothetical outer tertiary orbit of Binary 7782.  The solid diagonal black lines show the circularization radius a$_{\rm c}$ for the mass transfer stream coming from the outer tertiary, for different values of the mas ratio, namely q $=$ 0.1, 0.5, 1, 2 and 10.  We assume initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and m$_{\rm 3}$ = 1.4 M$_{\rm \odot}$ for the outer tertiary.  We assume completely conservative mass transfer for this exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal black line shows a rough criterion for dynamical stability in the triple, roughly following \citet{mardling99} (i.e., a$_{\rm in} <$ 0.1a$_{\rm out}$ is required).  The vertical solid red lines show the outer semi-major axes corresponding to the hard-soft boundary assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km s$^{-1}$.  The vertical dashed black lines show the maximum outer semi-major axis a$_{\rm out}$ for which the outer tertiary companion is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$ 200 R$_{\odot}$ (CHANGE).  The horizontal solid red line shows the semi-major axis corresponding to a contact state for the inner binary, assuming R$_{\rm 1} =$ R$_{\rm 2} =$ 1 R$_{\odot}$.  The horizontal dashed red line shows the observed semi-major axis for Binary 7782, for our assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2} =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line shows the parameter space for a$_{\rm out}$ allowed after considering all of the aforementioned criteria.
\label{fig:fig1}}
\end{figure}

@@

\item For the outer tertiary orbit, the semi-major axis should lie between 2 AU $\lesssim$ a$_{\rm out}$ $\lesssim$ 5 AU, assuming initial masses for the inner binary components of m$_{\rm 1} =$ 1.1 M$_{\odot}$ and m$_{\rm 2} =$ 0.9 M$_{\odot}$ and an outer tertiary mass of m$_{\rm 3} =$ 1.4 M$_{\odot}$.

@@

\section{Acknowledgments}

N.W.C.L. acknowledges support from a Kalbfleisch Fellowship at the American Museum of Natural History.  T.H.P. acknowledges the support through the FONDECYT Regular Project No. 1161817 and the BASAL Center for Astrophysics and Associated Technologies (PFB-06).

\begin{thebibliography}{99}

\bibitem[\protect\citeauthoryear{Anderson et al.}{2008}]{anderson08}
  Anderson J.,  Sarajedini A., Bedin L. R., King I. R., Piotto G.,
  Reid I. N., Siegel M., Majewski S. R., Paust N. E. Q., Aparicio A.,
  Milone A. P., Chaboyer B., Rosenberg A. 2008, AJ, 135, 2055
\bibitem[\protect\citeauthoryear{Artymowicz et al.}{1993}]{artymowicz93} Artymowicz P., Lin D. N. C., Wampler E. J. 1993, ApJ, 409, 592 

@@

\bibitem[\protect\citeauthoryear{Bedin et al.}{2010}]{bedin10} Bedin L. R., Cassisi S., Castelli F., Piotto G., Anderson J., Salaris M., Momany Y., Pietrinferni A. 2010, MNRAS 357, 1038

@@

\bibitem[\protect\citeauthoryear{Bruzual \& Charlot}{2003}]{bruzual03} Bruzual G., Charlot S. 2003, MNRAS, 344, 1000 
\bibitem[\protect\citeauthoryear{Casassus et al.}{2013}]{casassus13} Casassus S., Hales A., de Gregorio I., Dent B., Belloche A., G\:usten R., M\'enard F., Hughes A. M., Wilner D., Salinas V. 2013, A\&A, 553, 64 
\bibitem[\protect\citeauthoryear{Ceillier et al.}{2017}]{ceillier17} Ceillier T., Tayar J., Mathur S., Salabert D., Garc\'ia R. A., Stello D., Pinsonneault M. H., van Sanders J., Beck P. G., Bloemen S. 2017, A\& A, 605, 111 

@@

\bibitem[\protect\citeauthoryear{Dotter et al.}{2010}]{dotter10} Dotter, A., Sarajedini A., Anderson J., Aparicio A., Bedin L. R., Chaboyer B., Majewski S., Marin-Franch A., Milone A., Paust N., Piotto G., Reid N., Rosenberg A., Siegel M. 2010, ApJ 708, 698 
\bibitem[\protect\citeauthoryear{Dalessandro et al.}{2013}]{dalessandro13} Dalessandro E., Ferraro F. R., Massari D., Lanzoni B., Miocchi R. P., Beccari G., Bellini A., Sills A., Sigurdsson S., Mucciarelli A., Lovisi L. 2013, ApJ, 778, 135
\bibitem[\protect\citeauthoryear{Eggleton}{1983}]{eggleton83} Eggleton P. P. 1983, ApJ, 268, 368 
\bibitem[\protect\citeauthoryear{Flock et al.}{2015}]{flock15} Flock M., Ruge J. P., Dzyurkevich N., Henning Th., Klahr H., Wolf S. 2015, A\&A, 574, 68
\bibitem[\protect\citeauthoryear{Ferraro et al.}{1997}]{ferraro97}
  Ferraro F. R., Paltrinieri B., Fusi Pecci F., Cacciari C., Dorman
  B., Rood R. T., Buonanno R., Corsi C. E., Burgarella D., Laget
  M. 1997, A\&A, 324, 915
\bibitem[\protect\citeauthoryear{Ferraro et al.}{1999}]{ferraro99}
  Ferraro F. R., Paltrinieri B., Rood R. T., Dorman B. 1999, ApJ, 522,
  983
\bibitem[\protect\citeauthoryear{Ferraro et al.}{2004}]{ferraro04}
  Ferraro F. R., Beccari G., Rood, R. T., Bellazzini M., Sills A.,
  Sabbi E. 2004, ApJ, 603, 127
\bibitem[\protect\citeauthoryear{Ferraro, Valenti \& Origlia}{2006}]{ferraro06} Ferraro F. R., Valenti E., Origlia L. 2006, ApJ, 649, 243
\bibitem[\protect\citeauthoryear{Ferraro et al.}{2009}]{ferraro09} Ferraro F. R., Beccari G., Dalessandro E., Lanzoni B., Sills A., Rood R. T., Pecci F. F., Karakas A. I., Miocchi P., Bovinelli S. 2009, Nature, 462, 1028  
